### 2020-08-06 Manuscript/kPtx_paper.tex

\documentclass[11pt]{article}
% Command line: pdflatex -shell-escape compulse.tex
\usepackage{geometry} 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{times}
\usepackage{bm}
\usepackage{fixltx2e}
\usepackage[outerbars]{changebar}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{color}
\usepackage{tabularx}
\usepackage{ulem}
\usepackage{verbatim}

\epstopdfsetup{suffix=} % to remove 'eps-to-pdf' suffix from converted images
%\usepackage{todonotes} % use option disable to hide all comments

\usepackage[sort&compress]{natbib}
\bibpunct{[}{]}{,}{n}{,}{,}

\usepackage[noend]{algpseudocode}

\usepackage{dsfont}
\usepackage{relsize}

\usepackage{todonotes}

\graphicspath{{Figure/}}

%changing the Eq. tag to use [] when numbering. use \eqref{label} to reference equations in text.
\makeatletter
  \def\tagform@#1{\maketag@@@{[#1]\@@italiccorr}}
\makeatother

\linespread{1.5}
%\setlength{\parindent}{0in}

% the following command can be used to mark changes made due to reviewers' concerns. Use as \revbox{Rx.y} for reviewer x, concern y.
\newif\ifmarkedup
\markeduptrue

\ifmarkedup
	\newcommand{\revbox}[1]{\marginpar{\framebox{\textcolor{blue}{#1}}}}
\else
	\newcommand{\revbox}[1]{}
	\renewcommand{\textcolor}[1]{}
	\renewcommand{\sout}[1]{}
\fi

%\newcommand{\bop}{$\vert B_1^+ \vert$}
\newcommand{\kt}{$k_\textrm{T}$}
\newcommand{\bmap}{$B_1^+$}
\newcommand{\mytilde}{\raise.17ex\hbox{$\scriptstyle\mathtt{\sim}$}}  % tilde symbol
\mathchardef\mhyphen="2D

\begin{document}

\title{Tailored Spiral In-Out Spectral-Spatial Water Suppression Pulses for 
Magnetic Resonance Spectroscopic Imaging}
\author{Jun Ma$^{1,3}$, Carrie Wismans$^{2}$, Zhipeng Cao$^{1}$, Dennis W.J. Klomp$^{2}$, Jannie P. Wijnen$^{2}$\\ and William A. Grissom$^{*1,3}$}
\maketitle
\begin{flushleft}
$^1$Vanderbilt University Institute of Imaging Science, Nashville, TN, United States\\
$^2$Department of Radiology, University Medical Centre Utrecht, Utrecht, the Netherlands \\   
$^3$Department of Biomedical Engineering, Vanderbilt University, Nashville, TN, United States\\

\par
--------------------------

\par
Word Count: Approximately 3550 \\
*Corresponding author: \\
Will Grissom\\
Department of Biomedical Engineering\\
Vanderbilt University\\
5824 Stevenson Center\\
Nashville, TN 37235 USA \\
E-mail: will.grissom@vanderbilt.edu \\
Twitter: @wgrissom

\par Submitted to Magnetic Resonance in Medicine for consideration as a Full Paper.

\par
Acknowledgment: This work was supported by NIH R01 EB016695.

\end{flushleft}
\thispagestyle{plain}

\pagebreak

%%%%%%%%%%%%%%%%%%%%%%% Abstract %%%%%%%%%%%%%%%%%%%%%%%

\section*{\underline{Abstract}} 
{\bf Purpose:}
something
\\[1em]
{\bf Methods:}
something
\\[1em]
{\bf Results:}
something
\\[1em]
{\bf Conclusion:}
something
\\[1em]
{\bf \noindent Key words:} Spectroscopy; RF pulses; Ultra-high field MRI; RF pulse design; Selective excitation; Optimization.

\pagebreak

%%%%%%%%%%%%%%%%%%%%%%% Main Text %%%%%%%%%%%%%%%%%%%%%%%

\input{Introduction}

\input{Theory}

\input{Method}

\input{Result}

\input{Conclusion}

\pagebreak

\bibliographystyle{unsrt}  % cse doesn't work
\bibliography{kPtx_paper}

\pagebreak

\end{document}