### 2020-09-10 Manuscript/kPtx_paper.tex

\documentclass[11pt]{article}
% Command line: pdflatex -shell-escape compulse.tex
\usepackage{geometry} 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{times}
\usepackage{bm}
\usepackage{fixltx2e}
\usepackage[outerbars]{changebar}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{color}
\usepackage{tabularx}
\usepackage{ulem}
\usepackage{verbatim}
\usepackage{textcomp}

\epstopdfsetup{suffix=} % to remove 'eps-to-pdf' suffix from converted images
%\usepackage{todonotes} % use option disable to hide all comments

\usepackage[sort&compress]{natbib}
\bibpunct{[}{]}{,}{n}{,}{,}

\usepackage[noend]{algpseudocode}

\usepackage{dsfont}
\usepackage{relsize}

\usepackage{todonotes}

\graphicspath{{Figure/}}

%changing the Eq. tag to use [] when numbering. use \eqref{label} to reference equations in text.
\makeatletter
  \def\tagform@#1{\maketag@@@{[#1]\@@italiccorr}}
\makeatother

\linespread{1.5}
%\setlength{\parindent}{0in}

% the following command can be used to mark changes made due to reviewers' concerns. Use as \revbox{Rx.y} for reviewer x, concern y.
\newif\ifmarkedup
\markeduptrue

\ifmarkedup
	\newcommand{\revbox}[1]{\marginpar{\framebox{\textcolor{blue}{#1}}}}
\else
	\newcommand{\revbox}[1]{}
	\renewcommand{\textcolor}[1]{}
	\renewcommand{\sout}[1]{}
\fi

%\newcommand{\bop}{$\vert B_1^+ \vert$}
\newcommand{\kt}{$k_\textrm{T}$}
\newcommand{\bmap}{$B_1^+$}
\newcommand{\mytilde}{\raise.17ex\hbox{$\scriptstyle\mathtt{\sim}$}}  % tilde symbol
\mathchardef\mhyphen="2D

\begin{document}

\title{k-Space Domain Parallel Transmit Pulse Design}
\author{Jun Ma$^{1,2}$, Bernhard Gruber$^{3,4}$, Xinqiang Yan$^{1,5}$, and William A. Grissom$^{1,2,5*}$}
\maketitle
\begin{flushleft}
\vspace{-0.5cm}
$^1$Vanderbilt University Institute of Imaging Science, Nashville, TN, United States\\
$^2$Department of Biomedical Engineering, Vanderbilt University, Nashville, TN, United States\\
$^3$A. A. Martinos Center for Biomedical Imaging, Massachusetts General Hospital, Harvard Medical School, Charlestown, MA, United States\\
$^4$Division MR Physics, Center for Medical Physics and Biomedical Engineering, Medical University Vienna, Vienna, Austria\\
$^5$Department of Radiology and Radiological Sciences, Vanderbilt University, Nashville, TN, United States\\

\par
--------------------------

\par
Word Count: Approximately 5500 \\
*Corresponding author: \\
Will Grissom\\
Department of Biomedical Engineering\\
Vanderbilt University\\
5824 Stevenson Center\\
Nashville, TN 37235 USA \\
E-mail: will.grissom@vanderbilt.edu \\
Twitter: @wgrissom

\par Submitted to Magnetic Resonance in Medicine for consideration as a Full Paper.

\par
Acknowledgment: This work was supported by NIH grants R01 EB016695 and U01 EB 025162.

\end{flushleft}
\thispagestyle{plain}

\pagebreak

%%%%%%%%%%%%%%%%%%%%%%% Abstract %%%%%%%%%%%%%%%%%%%%%%%

\section*{\underline{Abstract}} 
{\bf Purpose:}
To accelerate the design of multidimensional parallel transmission pulses. 
\\[1em]
{\bf Methods:}
A k-space domain parallel transmission pulse design algorithm was proposed that
produces a sparse matrix relating a target excitation pattern to the pulses that produce it,
and can be finely parallelized. 
The algorithm was applied in simulations to the design of 3D SPINS pulses for inner volume excitation in the brain at 7 Tesla.
It was characterized in terms of the dependence of computation time, excitation error, and required memory
on algorithm parameters,
and it was compared to the spatial domain pulse design method in terms computation time, excitation error,
Gibbs ringing, and ability to compensate off-resonance.
\\[1em]
{\bf Results:}
The proposed algorithm achieved approximately 90\% faster pulse design compared to 
an iterative spatial domain method, with the same number of parallel threads,
with the tradeoff of increased excitation error and RMS RF amplitude. 
It reduced the memory required to store the design matrix by 98\% compared to a full matrix solution.
Even with a coarse design grid, the algorithm produced patterns that were free of Gibbs ringing.
It was similarly sensitive to k-space undersampling as the spatial domain method,
and was similarly capable of compensating for off-resonance.
\\[1em]
{\bf Conclusion:}
The proposed k-space domain algorithm accelerates and finely parallelizes parallel transmission pulse design,
with a modest tradeoff of excitation error and RMS RF amplitude.
\\[1em]
{\bf \noindent Key words:} Parallel Transmission; RF pulses; Ultra-high field MRI; RF pulse design; Selective excitation.

\pagebreak

%%%%%%%%%%%%%%%%%%%%%%% Main Text %%%%%%%%%%%%%%%%%%%%%%%

\input{Introduction}

\input{Theory}

\input{Method}

\input{Result}

\input{Conclusion}

\pagebreak

\bibliographystyle{cse}  % cse doesn't work
\bibliography{kPtx_paper}

\pagebreak
%(Please see main document for caption with intended math typesetting)

\begin{figure}
	\centering
	%\includegraphics[width=\textwidth]{kspace_PTX_Patch}
	\caption{(a) A target point $\vec{k}_j$ and the nearby trajectory points that are included when calculating 
	RF pulse weight contributions from this point.
	Inclusion width is defined as the distance from each target location $\vec{k}_j$ within which 
	which all trajectory points are considered when solving for the $j$-th column of the design ($\bm{W}$) matrix.
	(b) Calculation of the $\bm{W}$ matrix can be performed column-by-column for each target location $\vec{k}_j$,
	or patches of target points can be solved for jointly. 
	Here the 16 columns of $\bm{W}$ for a 4 $\times$ 4 patch of target points are solved together,
	and an inclusion width of 4 cycles/FOV dictates that all trajectory points within a 12 $\times$ 12 region
	centered on the patch are considered in the calculation.}
	\label{fig:Patch}
\end{figure}

\begin{figure}
	\centering
	%\includegraphics[width=6cm]{kspace_PTX_Coil}
	\caption{The 24-channel loop Tx array that was simulated in a human head model to obtain $B_1^+$ maps. 
	The array has diameter 32 cm and height 28 cm. The 16 cm $\times$ 11 cm rectangular loops are arranged in 3 rows of 8.}
	\label{fig:Coil}
\end{figure}

\begin{figure}
	\centering
	%\includegraphics[width=\textwidth]{kSpace_PTX_Pattern_Trajectory}
	\caption{(a) Middle axial, sagittal, and coronal slices of the target excitation pattern used for all pulse designs. 
	(b) The SPINS trajectory used in the designs. 
	(c) 10 ms minimum-time gradient waveforms that produce the SPINS trajectory.}
	\label{fig:Target}
\end{figure}

\begin{figure}
	\centering
	%\includegraphics[width=\textwidth]{ErrorMap}
	\caption{ %\textcolor{red}{WAG: I think this is the figure that your example in Github should replicate.}
	Normalized excitation patterns (top row) and error maps (bottom row) in central axial, sagittal and coronal slices 
	for k-space domain (left) and spatial domain designs (right).}
	\label{fig:ErrorMap}
\end{figure}

\begin{figure}
	\centering
	%\includegraphics[width=\textwidth]{ComputationTime}
	\caption{(a) k-Space computation time versus number of parallel threads, for patch and inclusion widths of 4 cycles/FOV. 
	(b) Computation time (blue axis) and RMSE (red axis) versus patch width, for an inclusion width of 4 cycles/FOV and 16 threads. 
	(c) Computation time (blue axis) and RMSE (red axis) versus inclusion width, for patch widths of 4 and 8 cycles/FOV, and 16 threads.}
	\label{fig:ComputationTime}
\end{figure}

\begin{figure}
\centering
%\begin{tabular}{c | c}
%Inclusion Width & $\bm{W}$ Matrix Size (GB) \\
%\hline
%$\infty$ & 24.4 \\
%8 & 2.73 \\
%6 & 1.37 \\
%4 & 0.53 \\
%2 & 0.14
%\end{tabular}
\caption{$\bm{W}$ matrix sizes in gigabytes (GB) versus inclusion width in cycles/FOV. 
An inclusion width of $\infty$ corresponds to a full matrix solution.}
\label{fig:wsize}
\end{figure}

\begin{figure}
	\centering
	%\includegraphics[width=\textwidth]{L_curve}
	\caption{L-curves for spatial domain and k-space domain pulse designs, 
	repeated across five orders of magnitude of the methods' Tikhonov regularization parameters.
	The k-space domain designs were repeated using patch and inclusion widths of four versus solving for the entire domain at once (`Patch / Inclusion Widths = 4' versus `Patch / Inclusion Widths = $\infty$'),
	and using $B_1^+$ map product interpolation versus phase modulation to each trajectory location (`Interpolated Matrices' versus
	`Exact Matrices').}
	\label{fig:LCurves}
\end{figure}

\begin{figure}
	\centering
	%\includegraphics[width=\textwidth]{GibbsRinging}
	\caption{Normalized error maps and RMSEs for 
	(a) pulses designed using a trajectory that reaches 6 mm isotropic resolution
	and the spatial domain algorithm with a 32$\times$32$\times$24 grid, 
	(b) pulses designed using the same trajectory and the spatial domain algorithm with a 64$\times$64$\times$48 grid,  
	and (c) pulses designed using the same trajectory and the k-space domain algorithm with a 32$\times$32$\times$24 design grid. 
	Red arrows indicate Gibbs ringing in the 32$\times$32$\times$24 spatial domain design.}
	\label{fig:GibbsRing}
\end{figure}

\begin{figure}
	\centering
	%\includegraphics[width=\textwidth]{ReductionFactor}
	\caption{Normalized error maps and RMSEs for spatial domain design (first row) and k-space domain design (second row), 
	using excitation k-space trajectories with different reduction factors (third row).
	The reduction factors are referenced to the 10 ms trajectory in Figure \ref{fig:Target} (second column).}
	\label{fig:kspace_PTX_Acceleration}
\end{figure}

\begin{figure}
	\centering
	%\includegraphics[width=\textwidth]{OffResonance}
	\caption{
	(a) Normalized off-resonance map containing a Gaussian distortion centered above the frontal sinus, 
	to mimic air-tissue susceptibility difference-induced $B_0$ inhomogeneity.
	(b) Normalized excitation maps and RMSEs for spatial domain design with off-resonance correction (first row), 
	and k-space domain design without and with off-resonance correction (second and third rows).}
	\label{fig:kspace_PTX_B0}
\end{figure}

\end{document}