### 2020-10-01 Manuscript_r1/coverletter/coverletter.tex

\documentclass[11pt]{article}

\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{marvosym}
\usepackage{fancyhdr}
\usepackage{ifpdf}
\ifpdf
  \usepackage[pdftex]{graphicx}
\else
  \usepackage[dvips]{graphicx}\fi

\textwidth 6.6in
%\textheight 9in
%\headheight 2in
\voffset 0.65in
\headheight 3in

\pagestyle{fancy}
\rhead{\includegraphics[width=3in]{vu02a.pdf}}
\chead{}
\lhead{\includegraphics[width=0.7in]{New_VUIIS_Logo.png}}
\lfoot{}
\cfoot{\small 1161 21st Ave South, AA-3114 MCN, Nashville, TN 37232-2310    Tel: (919) 649 4699  Fax: (615) 322 0734}
\rfoot{}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}

\usepackage[scale=0.775]{geometry}
\setlength{\parindent}{0pt}
\addtolength{\parskip}{6pt}

\def\firstname{William A.}
\def\familyname{Grissom, Ph.D.}
\def\FileAuthor{\firstname \familyname}
\def\FileTitle{\firstname \familyname's Cover Letter}
\def\FileSubject{Support Letter}
\def\FileKeyWords{\firstname \familyname, Cover Letter}

\renewcommand{\ttdefault}{pcr}

\usepackage{url}
\urlstyle{tt}
\ifpdf
  \usepackage[pdftex,pdfborder=0,breaklinks,baseurl=http://,pdfpagemode=None,pdfstartview=XYZ,pdfstartpage=1]{hyperref}
  \hypersetup{
    pdfauthor   = \FileAuthor,%
    pdftitle    = \FileTitle,%
    pdfsubject  = \FileSubject,%
    pdfkeywords = \FileKeyWords,%
    pdfcreator  = \LaTeX,%
    pdfproducer = \LaTeX}
\else
  \usepackage[dvips]{hyperref}
\fi

\begin{document}
\sffamily   % for use with a résumé using sans serif fonts;
%\rmfamily  % for use with a résumé using serif fonts;
%\hfill%
%\begin{minipage}[t]{.6\textwidth}
\raggedright%
{\bfseries \firstname~\familyname}\\[.35ex]
\small{\itshape%
Associate Professor of Biomedical Engineering\\
Institute of Imaging Science\\
Vanderbilt University\\}
\Letter~\href{mailto:will.grissom@vanderbilt.edu}{will.grissom@vanderbilt.edu}\\[2em]
{\bfseries Prof. Peter Jezzard}\\[.35ex]
\small{\itshape
Editor in Chief\\
Magnetic Resonance in Medicine\\[1em]}
%\end{minipage}
%
%\begin{minipage}[t]{.51\textwidth}
%\raggedright%
%{\bfseries Chair of Graduate Admissions \\ 
%Center for Biomedical Engineering \\
%University of Kentucky}\\[.35ex]
%\end{minipage}
%\hfill % US style
%\\[1em] % UK style
\large
\hfill\begin{minipage}[t]{\textwidth} % was 0.4\textwidth when addressee was there
\raggedleft % US style
\today
%April 6, 2006 % US informal style
%05/04/2006 % UK formal style
\end{minipage}\\[2em]
\raggedright
Dear Dr. Jezzard,\\[1.5em]
%
I would like to submit the manuscript `k-Space Domain Parallel Transmit Pulse Design' to be considered
for publication as a Full Paper in Magnetic Resonance in Medicine. 
It describes a new k-space-based algorithm for designing parallel transmission pulses that can be finely parallelized
and may be especially useful for designing $>$ 2D dimensional pulses with large numbers of transmit channels
for ultra-high field MRI.

\par Thank you very much for considering this manuscript. 
Please do not hesitate to contact me if you have any questions.\\[1.5em]
% 
%Yours sincerely,\\[2em] % if the opening is "Dear Mr(s) Doe,"
Sincerely,\\[1.5em] % if the opening is "Dear Sir or Madam,"
%
%\vspace{3cm}
%\includegraphics[scale=0.2]{/Users/guille/Documents/signature.png}\\[1.5em]
{\bfseries Will Grissom}\\
%
\vfill%
%{\slshape Enclosure}

\end{document}