### 2020-10-23 Manuscript_r2/kPtx_paper.tex

\documentclass[11pt]{article}
% Command line: pdflatex -shell-escape compulse.tex
\usepackage{geometry} 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{times}
\usepackage{bm}
\usepackage{fixltx2e}
\usepackage[outerbars]{changebar}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{color}
\usepackage{tabularx}
\usepackage{ulem}
%\usepackage{verbatim}
\usepackage{textcomp}
\usepackage{hyperref}
\hypersetup{colorlinks=false,urlcolor=blue,linkcolor=blue}
\usepackage{caption}

\epstopdfsetup{suffix=} % to remove 'eps-to-pdf' suffix from converted images
%\usepackage{todonotes} % use option disable to hide all comments

\usepackage[sort&compress]{natbib}
\bibpunct{[}{]}{,}{n}{,}{,}

%\usepackage[noend]{algpseudocode}

\usepackage{dsfont}
\usepackage{relsize}

%\usepackage{todonotes}

%\graphicspath{{Figure/}}

%changing the Eq. tag to use [] when numbering. use \eqref{label} to reference equations in text.
\makeatletter
  \def\tagform@#1{\maketag@@@{[#1]\@@italiccorr}}
\makeatother

\linespread{1.5}
%\setlength{\parindent}{0in}

% the following command can be used to mark changes made due to reviewers' concerns. Use as \revbox{Rx.y} for reviewer x, concern y.
\newif\ifmarkedup
\markeduptrue

\ifmarkedup
	\newcommand{\revbox}[1]{\marginpar{\framebox{\textcolor{blue}{#1}}}}
\else
	\newcommand{\revbox}[1]{}
	\renewcommand{\textcolor}[1]{}
	\renewcommand{\sout}[1]{}
\fi

%\newcommand{\bop}{$\vert B_1^+ \vert$}
\newcommand{\kt}{$k_\textrm{T}$}
\newcommand{\bmap}{$B_1^+$}
\newcommand{\mytilde}{\raise.17ex\hbox{$\scriptstyle\mathtt{\sim}$}}  % tilde symbol
\mathchardef\mhyphen="2D

\begin{document}

\title{k-Space Domain Parallel Transmit Pulse Design}
\author{Jun Ma$^{1,2}$, Bernhard Gruber$^{3,4}$, Xinqiang Yan$^{1,5}$, and William A. Grissom$^{1,2,5*}$}
\maketitle
\begin{flushleft}
\vspace{-0.5cm}
$^1$Vanderbilt University Institute of Imaging Science, Nashville, TN, United States\\
$^2$Department of Biomedical Engineering, Vanderbilt University, Nashville, TN, United States\\
$^3$A. A. Martinos Center for Biomedical Imaging, Massachusetts General Hospital, Harvard Medical School, Charlestown, MA, United States\\
$^4$Division MR Physics, Center for Medical Physics and Biomedical Engineering, Medical University Vienna, Vienna, Austria\\
$^5$Department of Radiology and Radiological Sciences, Vanderbilt University, Nashville, TN, United States\\

\par
--------------------------

\par
Word Count: 5630 \\
*Corresponding author: \\
Will Grissom\\
Department of Biomedical Engineering\\
Vanderbilt University\\
5824 Stevenson Center\\
Nashville, TN 37235 USA \\
E-mail: will.grissom@vanderbilt.edu \\
Twitter: @wgrissom

\par Submitted to Magnetic Resonance in Medicine for consideration as a Full Paper.

\par
Acknowledgment: This work was supported by NIH grants R01 EB016695 and U01 EB 025162.

\end{flushleft}
\thispagestyle{plain}

\pagebreak

%%%%%%%%%%%%%%%%%%%%%%% Abstract %%%%%%%%%%%%%%%%%%%%%%%

\section*{\underline{Abstract}} 
{\bf Purpose:}
To accelerate the design of \textcolor{blue}{(under- or oversampled)}\revbox{R1.4} multidimensional parallel transmission pulses. 
\\[1em]
{\bf Methods:}
A k-space domain parallel transmission pulse design algorithm was proposed that
produces a sparse matrix relating a \textcolor{blue}{complex-valued}\revbox{R1.1} target excitation pattern to the pulses that produce it, 
and can be finely parallelized. 
The algorithm was applied in simulations to the design of 3D SPINS pulses for inner volume excitation in the brain at 7 Tesla.
It was characterized in terms of the dependence of computation time, excitation error, and required memory
on algorithm parameters,
and it was compared to the spatial domain pulse design method in terms of computation time, excitation error,
Gibbs ringing, and ability to compensate off-resonance.
\\[1em]
{\bf Results:}
The proposed algorithm achieved approximately 80\% faster pulse design compared to 
an iterative spatial domain method, with the same number of parallel threads,
with the tradeoff of increased excitation error and RMS RF amplitude. 
It reduced the memory required to store the design matrix by 99\% compared to a full matrix solution.
Even with a coarse design grid, the algorithm produced patterns that were free of Gibbs ringing.
It was similarly sensitive to k-space undersampling as the spatial domain method,
and was similarly capable of compensating for off-resonance.
\\[1em]
{\bf Conclusion:}
The proposed k-space domain algorithm accelerates and finely parallelizes parallel transmission pulse design,
with a modest tradeoff of excitation error and RMS RF amplitude.
\\[1em]
{\bf \noindent Key words:} Parallel Transmission; RF pulses; Ultra-high field MRI; RF pulse design; Selective excitation.

\pagebreak

%%%%%%%%%%%%%%%%%%%%%%% Main Text %%%%%%%%%%%%%%%%%%%%%%%

\input{Introduction}

\input{Theory}

\input{Method}

\input{Result}

\input{Conclusion}

\pagebreak

\bibliographystyle{cse}
%\bibliography{kPtx_paper}

\begin{thebibliography}{10}
\providecommand{\url}[1]{\texttt{#1}}
\providecommand{\urlprefix}{URL }

\bibitem{Katscher:2003:Magn-Reson-Med:12509830}
Katscher U, B{\"o}rnert P, Leussler C, van~den Brink JS.
\newblock Transmit {SENSE}.
\newblock Magn Reson Med 2003;\hspace{0pt}49:144--150.

\bibitem{zhu2004parallel}
Zhu Y.
\newblock Parallel excitation with an array of transmit coils.
\newblock Magn Reson Med 2004;\hspace{0pt}51:775--784.

\bibitem{malik:mrm:2015}
Malik SJ, Hajnal JV.
\newblock Phase relaxed localized excitation pulses for inner volume fast spin
  echo imaging.
\newblock Magn Reson Med 2016;\hspace{0pt}76:848--861.

\bibitem{mooiweer:2016}
Mooiweer R, Sbrizzi A, Raaijmakers AJE, Van~den Berg CAT, Luijten PR, Hoogduin
  H.
\newblock Combining a reduced field of excitation with {SENSE}-based parallel
  imaging for maximum imaging efficiency.
\newblock Magn Reson Med 2016;\hspace{0pt}78:88--96.

\bibitem{Zhang:2007:Magn-Reson-Med:16526012}
Zhang Z, Yip CY, Grissom W, Noll DC, Boada FE, Stenger VA.
\newblock Reduction of transmitter {B1} inhomogeneity with transmit {SENSE}
  slice-select pulses.
\newblock Magn Reson Med 2007;\hspace{0pt}57:842--847.

\bibitem{cloos:kpstd:2012}
Cloos MA, Boulant N, Luong M, Ferrand G, Giacomini E, Le~Bihan D, Amadon A.
\newblock {$k_T$}-{Points}: {Short} three-dimensional tailored {RF} pulses for
  flip-angle homogenization over an extended volume.
\newblock Magn Reson Med 2011;\hspace{0pt}67:72--80.

\bibitem{deng:2009}
Deng W, Yang C, Alagappan V, Wald LL, Boada FE, Stenger VA.
\newblock Simultaneous z-shim method for reducing susceptibility artifacts with
  multiple transmitters.
\newblock Magn Reson Med 2009;\hspace{0pt}61:255--259.

\bibitem{Grissom:2006:MRM}
Grissom WA, Yip CY, Zhang Z, Stenger VA, Fessler JA, Noll DC.
\newblock Spatial domain method for the design of {RF} pulses in multicoil
  parallel excitation.
\newblock Magn Reson Med 2006;\hspace{0pt}56:620--9.

\bibitem{padormo:2016}
F.~Padormo, A.~Beqiri, J.~V. Hajnal, and S.~J. Malik.
\newblock Parallel transmission for ultrahigh-field imaging.
\newblock {NMR} Biomed 2016; 29(9):1145--1161.

\bibitem{deniz:2019}
C.~M. Deniz.
\newblock Parallel transmission for ultrahigh field {MRI}.
\newblock Top Magn Reson Imaging 2019; 28(3):159--171.

\bibitem{malik2012tailored}
Malik SJ, Keihaninejad S, Hammers A, Hajnal JV.
\newblock Tailored excitation in {3D} with spiral nonselective {(SPINS) RF}
  pulses.
\newblock Magn Reson Med 2012;\hspace{0pt}67:1303--1315.

\bibitem{davids2016fast}
Davids M, Schad LR, Wald LL, Guerin B.
\newblock Fast three-dimensional inner volume excitations using parallel
  transmission and optimized k-space trajectories.
\newblock Magn Reson Med 2016;\hspace{0pt}76:1170--1182.

\bibitem{setsompop2009}
K.~Setsompop, V.~Alagappan, B.~A. Gagoski, A.~Potthast, F.~Hebrank, U.~Fontius,
  F.~Schmitt, L.~L. Wald, and E.~Adalsteinsson.
\newblock Broadband slab selection with $b_1^+$ mitigation at 7{T} via parallel
  spectral-spatial excitation.
\newblock Magn Reson Med 2009; 61:493--500.

\bibitem{Malik:2010aa}
S.~J. Malik, D.~J. Larkman, D.~P. O'Regan, and J.~V. Hajnal.
\newblock Subject-specific water-selective imaging using parallel transmission.
\newblock Magn Reson Med 2010; 63(4):988--997.

\bibitem{stenger2000three}
Stenger VA, Boada FE, Noll DC.
\newblock Three-dimensional tailored {RF} pulses for the reduction of
  susceptibility artifacts in {$T_2^*$}-weighted functional {MRI}.
\newblock Magn Reson Med 2000;\hspace{0pt}44:525--531.

\bibitem{yang2010four}
Yang C, Deng W, Alagappan V, Wald LL, Stenger VA.
\newblock Four-dimensional spectral-spatial {RF} pulses for simultaneous
  correction of {B1+} inhomogeneity and susceptibility artifacts in
  {T2*}-weighted {MRI}.
\newblock Magn Reson Med 2010;\hspace{0pt}64:1--8.

\bibitem{orzada:2019}
Orzada S, Solbach K, Gratz M, Brunheim S, Fiedler TM, Johst S, Bitz AK,
  Shooshtary S, Abuelhaija A, Voelker MN, Rietsch SHG, Kraff O, Maderwald S,
  Fl\"oser M, Oehmigen M, Quick HH, Ladd ME.
\newblock A 32-channel parallel transmit system add-on for {7T MRI}.
\newblock PLOS ONE 2019;\hspace{0pt}14:1--20.

\bibitem{deng:2011}
Deng W, Yang C, Stenger VA.
\newblock Accelerated multidimensional radiofrequency pulse design for parallel
  transmission using concurrent computation on multiple graphics processing
  units.
\newblock Magn Reson Med 2011;\hspace{0pt}65:363--369.

\bibitem{setsompop2008magnitude}
Setsompop K, Wald L, Alagappan V, Gagoski B, Adalsteinsson E.
\newblock Magnitude least squares optimization for parallel radio frequency
  excitation design demonstrated at 7 {Tesla} with eight channels.
\newblock Magn Reson Med 2008;\hspace{0pt}59:908--915.

\bibitem{brunner2010optimal}
Brunner DO, Pruessmann KP.
\newblock Optimal design of multiple-channel {RF} pulses under strict power and
  {SAR} constraints.
\newblock Magn Reson Med 2010;\hspace{0pt}63:1280--1291.

\bibitem{hoyos:tmi:2014}
Hoyos-Idrobo A, Weiss P, Massire A, Amadon A, Boulant N.
\newblock On variant strategies to solve the magnitude least squares
  optimization problem in parallel transmission pulse design under strict {SAR}
  and power constraints.
\newblock IEEE Trans Med Imag 2014;\hspace{0pt}33:739--748.

\bibitem{grissom:ismrm18}
Grissom WA.
\newblock {k-Space} domain parallel transmit pulse design.
\newblock In Proceedings 26th Scientific Meeting, International Society for
  Magnetic Resonance in Medicine, Paris. 2018;\hspace{0pt} p. 3396.

\bibitem{luo2019grappa}
Luo T, Noll DC, Fessler JA, Nielsen JF.
\newblock A {GRAPPA} algorithm for arbitrary {2D/3D} non-{Cartesian} sampling
  trajectories with rapid calibration.
\newblock Magn Reson Med 2019;\hspace{0pt}82:1101--1112.

\bibitem{fessler2005toeplitz}
Fessler JA, Lee S, Olafsson VT, Shi HR, Noll DC.
\newblock Toeplitz-based iterative image reconstruction for {MRI} with
  correction for magnetic field inhomogeneity.
\newblock IEEE Trans Sig Proc 2005;\hspace{0pt}53:3393--3402.

\bibitem{lustig2008fast}
Lustig M, Kim SJ, Pauly JM.
\newblock A fast method for designing time-optimal gradient waveforms for
  arbitrary k-space trajectories.
\newblock {IEEE} Trans Med Imag 2008;\hspace{0pt}27:866--873.

\bibitem{Fessler:2003fk}
Fessler JA, Sutton BP.
\newblock Nonuniform fast {Fourier} transforms using min-max interpolation.
\newblock IEEE Trans Sig Proc 2003;\hspace{0pt}51:560--574.

\bibitem{Yip:2005:Magn-Reson-Med:16155881}
Yip CY, Fessler JA, Noll DC.
\newblock Iterative {RF} pulse design for multidimensional, small-tip-angle
  selective excitation.
\newblock Magn Reson Med 2005;\hspace{0pt}54:908--917.

\bibitem{cao2016array}
Cao Z, Yan X, Grissom WA.
\newblock Array-compressed parallel transmit pulse design.
\newblock Magn Reson Med 2016;\hspace{0pt}76:1158--1169.

\bibitem{Lee:2011:MRM}
D.~Lee, W.~A. Grissom, M.~Lustig, A.~B. Kerr, P.~P. Stang, and J.~M. Pauly.
\newblock {VERSE}-guided numerical {RF} pulse design: A fast method for peak
  {RF} power control.
\newblock Magn Reson Med 2012; 67:353--362.

\bibitem{lee2009tod}
D.~Lee, M.~Lustig, W.~A. Grissom, and J.~M. Pauly.
\newblock Time-optimal design for multidimensional and parallel transmit
  variable-rate selective excitation.
\newblock Magn Reson Med 2009, 61(6):1471--1479.

\bibitem{Xu:2008aa}
D.~Xu, K.~F. King, Y.~Zhu, G.~C. McKinnon, and Z.-P. Liang.
\newblock Designing multichannel, multidimensional, arbitrary flip angle {RF}
  pulses using an optimal control approach.
\newblock Magn Reson Med 2008, 59(3):547--560.

\bibitem{grissom:mrm:2008}
W.~A. Grissom, C.~Y. Yip, S.~M. Wright, J.~A. Fessler, and D.~C. Noll.
\newblock Additive angle method for fast large-tip-angle {RF} pulse design in
  parallel excitation.
\newblock Magn Reson Med 2008, 59(4):779--87.

\end{thebibliography}

\pagebreak

\section*{Tables}

\begin{table}[!h]
\centering
\begin{tabular}{c | c}
Inclusion Width & $\bm{W}$ Matrix Size (GB) \\
\hline
$\infty$ & 78.7 \\
8 & 4.09 \\
6 & 2.04 \\
4 & 0.79 \\
2 & 0.21
\end{tabular}
\caption{$\bm{W}$ matrix sizes in gigabytes (GB) versus inclusion width in cycles/FOV. 
An inclusion width of $\infty$ corresponds to a full matrix solution.}
\label{fig:wsize}
\end{table}

\pagebreak
%(Please see main document for caption with intended math typesetting)

\section*{Figure Captions}

\begin{figure}[!h]
	\centering
	%\includegraphics[width=\textwidth]{kspace_PTX_Patch}
	\caption{(a) A target point $\vec{k}_j$ and the nearby trajectory points that are included when calculating 
	RF pulse weight contributions from this point.
	Inclusion width is defined as the distance from each target location $\vec{k}_j$ within which all trajectory points are considered when solving for the $j$-th column of the design ($\bm{W}$) matrix.
	(b) Calculation of the $\bm{W}$ matrix can be performed column-by-column for each target location $\vec{k}_j$,
	or patches of target points can be solved for jointly. 
	Here the 16 columns of $\bm{W}$ for a 4 $\times$ 4 patch of target points are solved together,
	and an inclusion width of 4 cycles/FOV dictates that all trajectory points within a 12 $\times$ 12 region
	centered on the patch are considered in the calculation.}
	\label{fig:Patch}
\end{figure}

\begin{figure}[!h]
	\centering
	%\includegraphics[width=6cm]{kspace_PTX_Coil}
	\caption{The 24-channel loop Tx array that was simulated in a human head model to obtain $B_1^+$ maps. 
	The array has diameter 32 cm and height 28 cm. The 16 cm $\times$ 11 cm rectangular loops are arranged in 3 rows of 8.}
	\label{fig:Coil}
\end{figure}

\begin{figure}[!h]
	\centering
	%\includegraphics[width=\textwidth]{kSpace_PTX_Pattern_Trajectory}
	\caption{(a) Middle axial, sagittal, and coronal slices of the target excitation pattern used for all pulse designs. 
	(b) The SPINS trajectory used in the designs. 
	(c) 10 ms minimum-time gradient waveforms that produce the SPINS trajectory.}
	\label{fig:Target}
\end{figure}

\begin{figure}[!h]
	\centering
	%\includegraphics[width=\textwidth]{ErrorMap}
	\caption{ %\textcolor{red}{WAG: I think this is the figure that your example in Github should replicate.}
	Normalized excitation patterns (top row) and error maps (bottom row) in central axial, sagittal and coronal slices 
	for k-space domain (left) and spatial domain designs (right).}
	\label{fig:ErrorMap}
\end{figure}

\begin{figure}[!h]
	\centering
	%\includegraphics[width=\textwidth]{ComputationTime}
	\caption{(a) k-Space computation time versus number of parallel threads, for patch and inclusion widths of 4 cycles/FOV. 
	(b) Computation time (blue axis) and RMSE (red axis) versus patch width, for an inclusion width of 4 cycles/FOV and 16 threads. 
	(c) Computation time (blue axis) and RMSE (red axis) versus inclusion width, for patch widths of 4 cycles/FOV, and 16 threads.}
	\label{fig:ComputationTime}
\end{figure}

\begin{figure}[!h]
	\centering
	%\includegraphics[width=\textwidth]{L_curve}
	\caption{L-curves for spatial domain and k-space domain pulse designs, 
	repeated across five orders of magnitude of the methods' Tikhonov regularization parameters.
	The k-space domain designs were repeated using patch and inclusion widths of four versus solving for the entire domain at once (`Patch / Inclusion Widths = 4' versus `Patch / Inclusion Widths = $\infty$'),
	and using $B_1^+$ map product interpolation versus phase modulation to each trajectory location (`Interpolated Matrices' versus
	`Exact Matrices').}
	\label{fig:LCurves}
\end{figure}

\begin{figure}[!h]
	\centering
	%\includegraphics[width=\textwidth]{GibbsRinging}
	\caption{Normalized error maps and RMSEs for 
	(a) pulses designed using a trajectory that reaches 6 mm isotropic resolution
	and the spatial domain algorithm with a 32$\times$32$\times$24 grid, 
	(b) pulses designed using the same trajectory and the spatial domain algorithm with a 64$\times$64$\times$48 grid,  
	and (c) pulses designed using the same trajectory and the k-space domain algorithm with a 32$\times$32$\times$24 design grid. 
	Red arrows indicate Gibbs ringing in the 32$\times$32$\times$24 spatial domain design.}
	\label{fig:GibbsRing}
\end{figure}

\begin{figure}[!h]
	\centering
	%\includegraphics[width=\textwidth]{ReductionFactor}
	\caption{Normalized error maps, RMSEs, and computation times 
	for spatial domain design (first row) and k-space domain design (second row), 
	using excitation k-space trajectories with different reduction factors (third row).
	The reduction factors are referenced to the 10 ms trajectory in Figure \ref{fig:Target} (second column).}
	\label{fig:kspace_PTX_Acceleration}
\end{figure}

\begin{figure}[!h]
	\centering
	%\includegraphics[width=\textwidth]{OffResonance}
	\caption{
	(a) Normalized off-resonance map containing a Gaussian distortion centered above the frontal sinus, 
	to mimic air-tissue susceptibility difference-induced $B_0$ inhomogeneity.
	(b) Normalized excitation maps and RMSEs for spatial domain design with off-resonance correction (first row), 
	and k-space domain design without and with off-resonance correction (second and third rows).}
	\label{fig:kspace_PTX_B0}
\end{figure}

\end{document}